{Gravitational wave sources in our Galactic backyard:}\\{Predictions for BHBH, BHNS and NSNS binaries detectable with LISA

### 2022-07-06 paper/tex/sections/conclusion.tex

We provide predictions for the detection rate and population properties of LISA detectable BHBH, BHNS and NSNS.
%
We use a novel empirically-informed analytical model for the metallicity dependent star formation history of the Milky Way, calibrated against the APOGEE stellar spectroscopic survey. We use this to model Monte-Carlo realisations of the present-day BHBH, BHNS and NSNS populations in our Milky Way. 
%
For the binary population, we use the results of a large grid of simulations performed with the rapid population synthesis code COMPAS \citep{Broekgaarden+2021,Broekgaarden+2021b}. These simulations have been optimised with the adaptive sampling algorithm STROOPWAFEL to preferentially sample NS and BH binaries. In total these comprise over two billion massive binaries that span 20 physics variations, which represent the most common uncertainties in binary physics.
%
To determine the detectability of sources with LISA we use the LEGWORK package \citep{Wagg+2021}, that was specifically developed for this purpose and is publicly available. 
%
We investigate the results expected for a 4- and 10-year LISA mission. Our main conclusions can be summarised as:
\begin{enumerate}
    \item \textbf{Total detections:} We predict \rangeFourYear{} detections in a 4-year LISA mission, across all our simulations for varying physics assumptions. This increases to about \rangeTenYear{} for a 10-year LISA mission. Although the number of detections per type can vary by about 2 orders of magnitude, we find that the total detection rate is fairly robust, among the variations we have considered (see Table~\ref{tab:detection_rates}).

@@

    \item \textbf{Probing the black hole mass distribution and the lower mass gap:} We expect LISA to predominantly detect lower mass BHs (with 90\% of BHBH and BHNSs having BH masses lower than $11 \unit{M_\odot}$ in our fiducial simulations) in stark contrast to current ground-based detectors which are heavily biased towards high mass systems. For our fiducial simulation, we predict that approximately 69 systems with a component with a mass between $2.5$-$5 \unit{M_\odot}$ would be detected by a 4-year LISA. This implies that LISA can potentially make important contributions to the debate about the existence of a lower mass gap (see Fig.~\ref{fig:lower_mass_gap_variation}).

>>>>

We provide predictions for the detection rate and population properties of LISA detectable BHBH, BHNS and NSNS.
%
We use a novel empirically-informed analytical model for the metallicity dependent star formation history of the Milky Way, calibrated against the APOGEE stellar spectroscopic survey. We use this to model Monte-Carlo realisations of the present-day BHBH, BHNS and NSNS populations in our Milky Way. 
%
For the binary population, we use the results of a large grid of simulations performed with the rapid population synthesis code COMPAS\edit1{, which can reproduce the current constraints on the overall GW rates \citep{Broekgaarden+2021,Broekgaarden+2021b, Abbott+2021_GWTC3}}. Our suite of simulations consider 20 physics variations, which represent the most common uncertainties in binary physics.
%
We investigate the detectability of sources in the planned space-based GW detector LISA using the LEGWORK package \citep{Wagg+2021}, that we specifically developed for this purpose and is publicly available. 
%
We investigate the results expected for a 4- and 10-year LISA mission. Our main conclusions can be summarised as:
\begin{enumerate}
    \item \textbf{Total detections:} We predict \rangeFourYear{} detections in a 4-year LISA mission, across all our simulations for varying physics assumptions. This increases to about \rangeTenYear{} for a 10-year LISA mission. Although the number of detections per type can vary by about 2 orders of magnitude, we find that the total detection rate is fairly robust, among the variations we have considered (see Table~\ref{tab:detection_rates}).

@@

    \item \textbf{Probing the black hole mass distribution and the lower mass gap:} We expect LISA to predominantly detect lower mass BHs (with 90\% of BHBH and BHNSs having BH masses lower than $11 \unit{M_\odot}$ in our fiducial simulations). This is in stark contrast to current ground-based detectors which are heavily biased towards high mass systems, since they can be detected out to larger distances. We predict that between 15-156 systems with a component with a mass between $2.5$-$5 \unit{M_\odot}$ would be detected by a 4-year LISA, for our different physics variations. An exception is model using the Fryer Rapid remnant mass prescription, which produces none. However, it will likely be hard for LISA data to make a significant impact on this debate on the existence of the lower mass gap given the uncertainty of the individual masses (see Fig.~\ref{fig:lower_mass_gap_variation}).

### 2022-07-06 paper/tex/sections/discussion.tex

\paragraph{Progenitor models} Our binary-star progenitors models have been computed with a rapid population synthesis code (see Sec.~\ref{sec:COMPAS_explained}). This code relies on approximate parametric prescriptions for the stellar evolutionary tracks of single tracks and simple algorithms to mimic the effects of evolutionary and binary interaction  processes. Even though we explicitly consider the impact of some of the main physics uncertainties (see Sect.~\ref{sec:variation_assumptions}) the list of variations that we considered is far from exhaustive. Moreover, it is by no means guaranteed that the parametric prescriptions used in this code lead to realistic results, even when varying the values of the parameters to their extremes. We stress in particular the uncertainties affecting our most massive progenitor models. Observational constraints are scarce for high mass stars and practically non-existent for the rapid evolutionary phases \citep[e.g.][]{Langer2012, Mapelli+2021}. This is even more true for the evolution of massive stars at low metallicity. In addition to our limited understanding of massive stars, we note that the rapid population synthesis code, such as the one employed to compute the models used in this study, rely on extrapolations of the original fitting formulae to approximate the evolutionary tracks for these higher mass progenitors \citep{Hurley+2000,Hurley+2002}.

>>>>

\paragraph{Progenitor models} Our binary-star progenitors models have been computed with a rapid population synthesis code (see Sec.~\ref{sec:COMPAS_explained}). This code relies on approximate parametric prescriptions for the stellar evolutionary tracks of single tracks and simple algorithms to mimic the effects of evolutionary and binary interaction  processes. Even though we explicitly consider the impact of some of the main physics uncertainties (see Sect.~\ref{sec:variation_assumptions}) the list of variations that we considered is far from exhaustive. Moreover, it is by no means guaranteed that the parametric prescriptions used in this code lead to realistic results, even when varying the values of the parameters to their extremes. We stress in particular the uncertainties affecting our most massive progenitor models. Observational constraints are scarce for high mass stars and practically non-existent for the rapid evolutionary phases \citep[e.g.][]{Langer2012, Mapelli+2021}. This is even more true for the evolution of massive stars at low metallicity. In addition to our limited understanding of massive stars, we note that the rapid population synthesis code, such as the one employed to compute the models used in this study, rely on extrapolations of the original fitting formulae to approximate the evolutionary tracks for these higher mass progenitors \citep{Hurley+2000,Hurley+2002}. \edit1{A further caveat is whether the population synthesis predictions used in our study realistically describe the population of compact objects. \citet{Broekgaarden+2021b} shows that none of the physics variations that we use can be excluded at present based on the overall constraints on the GW rates, but this does not mean that they are accurate. For example, at present it is unclear whether these models reproduce the features in the inferred mass distribution \citep[van Son et al.\ in prep.][]{Farah+2021,TheLIGOScientificCollaboration+2021,Li+2021,Veske+2021,Tiwari2022,Edelman+2022}.}
